---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

The first thing to do is to load the libraries that we will need for data manipulation and graphs: in this case dplyr and ggplot2.  
We read the csv provided for this assignment through a read.csv, indicating also the presence of header and NA.
We then modify the "date" variable imported as a character in date.

```{R, message = FALSE}
library(dplyr)
library(ggplot2)
dati = read.csv("activity.csv", header = TRUE, sep = ",", na.strings = "NA")
dati$date = as.Date(as.character(dati$date))
```

## What is mean total number of steps taken per day?

Through dpylr pipes we omit the NAs, group by date and calculate number of steps taken per day.

```{r,message = FALSE}  
steps.per.day = dati %>%
  na.omit() %>%
  group_by(date) %>%
  summarise(total = sum(steps))
```
  
We visualize what has just been calculated through a histogram that we call "Steps per day".

```{r}  
par(mar = c(5,5,5,1))
hist(steps.per.day$total, col="#69b3a2", main = "Steps per day", xlab = "steps")
axis(2, at = seq(0,30,5), labels = seq(0,30,5))
```

The mean and the median of steps taken per day follow:

```{r}
mean(steps.per.day$total)
median(steps.per.day$total)
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
